`{traveltime}`: an R package to calculate walking and motorised travel time from user-specified locations

### 2025-01-29 paper/paper.qmd
alternate label and caption method for cross refs - fails in html render but will it work in yml pdf render?

Understanding and mapping the time to travel among locations is useful for many activities from urban planning [@zahavi1974traveltime] to public health [hulland2019travel; @weiss2020global] (@ref-tt_healthcare) and a myriad others [@nelson2019suite]. Here we present a software package --- `{traveltime}` --- written in and for the language R [@Rref]. `{traveltime}` enables a user to supply a set of geographic point locations and an extent of interest, to then calculate the motorised or walking travel time over that extent. The result is a raster of 30 arcsecond resolution (approximately 0.008333 decimal degrees, or just below 1 km$^2$ at the equator) where the value in each cell is the lowest travel time in minutes to any of the supplied point locations over the extent of interest (e.g., travel time from health care facilities @ref-tt_healthcare from @weiss2020global).

@@

```{r ttexample_image, echo=FALSE}
#| label: "fig-tt_healthcare"
#| fig-cap: "Existing surface of motorised travel time from healthcare facilities, exerpted to Cambodia (Weiss et al., 2020)"
library(ggplot2)
library(tidyterra)

@@

@weiss2020global and colleagues made their code available as an R script to allow for reproduction and extension of their analyses (<https://malariaatlas.org/wp-content/uploads/2022/11/R_generic_accessibilty_mapping_script_2020-1.txt> ). To further enable extension of this work, here we have developed an R package based on that code to seamlessly calculate the travel time from any arbitrary set of locations.

>>>>

Understanding and mapping the time to travel among locations is useful for many activities from urban planning [@zahavi1974traveltime] to public health (@hulland2019travel; @weiss2020global; \ref{fig:fig-tt_healthcare}) and a myriad others [@nelson2019suite]. Here we present a software package --- `{traveltime}` --- written in and for the language R [@Rref]. `{traveltime}` enables a user to supply a set of geographic point locations and an extent of interest, to then calculate the motorised or walking travel time over that extent. The result is a raster of 30 arcsecond resolution (approximately 0.008333 decimal degrees, or just below 1 km$^2$ at the equator) where the value in each cell is the lowest travel time in minutes to any of the supplied point locations over the extent of interest (e.g., travel time from health care facilities  \ref{fig:fig-tt_healthcare} from @weiss2020global).

@@

```{r ttexample_image, echo=FALSE, fig.cap = "\\label{fig:fig-tt_healthcare}Existing surface of motorised travel time from healthcare facilities, exerpted to Cambodia (Weiss et al., 2020)"}
library(ggplot2)
library(tidyterra)

@@

@weiss2020globa made their code available as an R script to allow for reproduction and extension of their analyses (<https://malariaatlas.org/wp-content/uploads/2022/11/R_generic_accessibilty_mapping_script_2020-1.txt> ). To further enable extension of this work, here we have developed an R package based on that code to seamlessly calculate the travel time from any arbitrary set of locations.
